Сплинты корневых систем для специальных подалгебр Ли

### 2015-03-23 AffineLieAlgebras/Splints/special-paper-ru.tex
More translation

The study of specialy embedded subalgebras traces back to the fundamental papers by Eugene Dynkin
\cite{dynkin1952semisimple,dynkin1952maximal} where he called such subalgebras ``S-subalgebras'' to
distinguish from regular or ``R-subalgebras'' that have root systems obtained by dropping some roots
of the algebra root system. Thus regular subalgebras are easy to construct by dropping nodes from
the extended Dynkin diagram of the algebra, while the case of special subalgebras is more difficult. The complete
classification for exceptional Lie algebras was obtained recently in \cite{minchenko2006semisimple}.
The algorithm of the special subalgebras construction is available in GAP package but still requires
a manual intervention \cite{de2011constructing}.

Assume that the Lie algebra $\gf$ is simple. Let's denote a subalgebra by
$\af$. We denote corresponding Cartan subalgebras by $\hfg$ and $\hfa$ and identify them with the dual
spaces $\hfg^{*}$, $\hfa^{*}$ using the Killing forms of $\gf$ and $\af$.

To construct an embedding $\af\to\gf$ consider some representation $L^{\nu}_{\af}$ as a subspace of
Lie algebra $\gf$. Then one needs to check that the generators of $\af$ can be presented as
linear combinations of generators of $\gf$. We can identify the Cartan subalgebra $\hfa\subset \af$ with
the dual space $\hfa^{*}$ using the Killing form. Then the root system $\Delta_{\gf}$ of $\gf$ can be
projected to $\hfa^{*}$ using the expression of $\hfa$-generators through $\hfg$-generators.

@@

This projection is described in the classical papers \cite{dynkin1952semisimple,dynkin1952maximal} by the following theorem:

\begin{theorem}\label{dyn0}
  If a representation $L^{\mu}_{\gf}$ of the algebra $\gf$ induces a representation
  $L^{\tilde\mu}_{\af}$ on the subalgebra $\af$ then the weight system of $L^{\tilde\mu}_{\af}$ can
  be obtained from $L^{\mu}_{\gf}$ by an orthogonal projection of $\hfg^{*}$ on $\hfa^{*}$.

  \cite{dynkin1952maximal}. 
\end{theorem}

In relation to the adjoint representation of $\gf$ this theorem means that the projection of the root system $\Delta_{\gf}$ to $\hfa^{*}$ is a weight system of some
finite-dimensional but not irreducible representation of $\af$. Moreover this reducible representation contains the adjoint representation of $\af$. And if we denote such a projection by
\begin{equation}
  \label{eq:2}
  \Delta'=\pia\left(\Delta_{\gf}\right),
\end{equation}
then the roots of $\af$ will be contained in $\Delta'$. Thus the system $\Delta'$ can be a root
system but it might not be reduced and might contain some vectors with multiplicities greater than
one.

The following theorems \cite{dynkin1952semisimple} also concern the properties of $\Delta'$.

\begin{theorem}\label{dyn1}
  Every special subalgebra $\af$ of a semisimple Lie algebra $\gf$ is integer, i.e. the projections of
  the roots of $\gf$ to $\hfa^{*}$ are linear combinations of the simple roots of $\af$ with integer
  coefficients \cite{dynkin1952semisimple}
\end{theorem}

\begin{theorem}\label{dyn2}
  If $\af$  is a semisimple subalgebra of a semisimple Lie algebra $\gf$ and the generator $e_{\alpha}$ corresponding to
  the root $\alpha$ of $\af$ is presented as a linear combination $e_{\alpha}=\sum_{\beta}
  e_{\beta}$ of $\gf$-generators
  corresponding to the roots $\beta$ of $\gf$, then the roots $\beta$ are projected into $\alpha$,
  $\pia(\beta)=\alpha$. 
  \cite{dynkin1972semisimple,dynkin1952semisimple}. 
\end{theorem}

From these theorems we see that the root system $\Delta_{\af}$ is metrically embedded (in the sense
of splint) into the projection $\Delta'$. Moreover, the multiplicities of some roots
$\alpha\in\Delta_{\af}$ in this projection $\Delta'$ are greater than one.

For example the projection of the root system of $D_{4}$($so(8)$) to the root system of the special
subalgebra $A_{2}$($su(3)$) coincides with the root system $G_{2}$, but the short roots have the multiplicity $3$.

@@

\section{Splints for special embeddings}
\label{sec:splints-spec-embedd}

@@

We want to classify all the cases when branching coefficients coincide with weight multiplicities of
some other algebra representations. Such a coincidence is possible when the projection $\Delta'$ of
the $\gf$ root system admits a splint $\Delta'=\varphi_{\af}(\Delta_{\af})\cup
\varphi_{\sfr}(\Delta_{\sfr})$, where $\varphi_{\af}$ and $\varphi_{\sfr}$ are the embeddings of
the corresponding root systems. The embedding $\varphi_{\af}$ is metric and trivial. Moreover, the
projection of the singular element $\pia\left(\Psi^{\mu}_{\gf}\right)$ should admit a decomposition into
a linear combination of singular elements of irreducible representations of $\sfr$:
\begin{equation}
  \label{eq:4}
  \pia\left(\Psi^{\mu}_{\gf}\right)=\sum_{\nu} \varkappa_{\nu}e^{\nu}\Psi^{\tilde\mu}_{\sfr},
\end{equation}
with integer coefficients $\varkappa_{\nu}$, where the sum is over the set of weights $\nu$ that
will be determined later. Such a decomposition of the singular element is very similar to the branching of Weyl group orbits considered in the papers \cite{larouche2011branching,larouche2009branching}. First we classify all the splints of the projection of a root system into
a union of simple root systems for special embeddings, and then discuss the decomposition of singular
elements.

Projection of the $\gf$ root system $\Delta'$ coincides with the projection of the weight diagram of the
adjoint representation of the algebra $\gf$ with the exception of zero weight. The adjoint representation of
$\gf$ contains the adjoint representation of $\af$ and can be decomposed as
\begin{equation}
  \label{eq:3}
  \mathrm{ad}_{\gf}=\mathrm{ad}_{\af}\oplus M_{\af}^{\chi},
\end{equation}
where $M^{\chi}_{\af}$ is not necessary irreducible representation of $\af$ called
``characteristic'' in \cite{dynkin1952semisimple}.

The weight diagram of $M^{\chi}_{\af}$ coincides with the root system $\Delta_{\sfr}$ with the exception of
zero weight. So we need to find all the representations for all simple Lie algebras $\af$, such that
their weight diagrams are root systems after the exclusion of zero weight. In order to do so we must
note, that all the weights of $M^{\chi}_{\af}$ should have a length not less than the length of the shortest
root of $\af$, since $\af$ is an integer subalgebra of $\gf$ \ref{dyn1}. Moreover, $M^{\chi}_{\af}$
should contain weights of at most two different lengths.

If the root system $\Delta_{\gf}$ contains roots $\alpha: \alpha\perp \beta,\; \forall \beta\in
\Delta_{\af},$ orthogonal to the root system of the subalgebra, one cannot proceed with the decomposition
of the singular element \ref{eq:4}. The elements $e^{\nu}$ of $\pia\left(\Psi^{\mu}_{\gf}\right)$
should be augmented with the dimensions of representations of the algebra $\afb$ spanned by generators
corresponding to the orthogonal roots \cite{2010arXiv1007.0318L}. Then there will be non-trivial
multiplicities in the right hand side of \ref{eq:4} and no coincidence of branching coefficients
with the weight multiplicities of $\sfr$-representations. One can write more cumbersome relation between
the multiplicities and branching coefficients, but it is out of scope for the present paper.

The multiplicity of zero weight in the adjoint representation is equal to the rank of the algebra. So if
$\mathrm{rank}\gf-\mathrm{rank}\af=1$, the representation in question $M^{\chi}_{\af}$ must be
multiplicity-free. If $\mathrm{rank}\gf-\mathrm{rank}\af>1$ only zero weight can have non-trivial
multiplicities.

The simplest class of multiplicity-free representations is called ``strongly multiplicity free''
\cite{lehrer2006strongly} and consists of multiplicity free representations with weight systems
admitting strict ordering $\nu_{1}<\nu_{2} \Leftrightarrow \nu_{1}=\nu_{2}+n \alpha$, where $n\in
\mathbb{N}, \alpha\in \Delta^{+}_{\af}$ and for all $\nu_{1},\nu_{2}$ either $\nu_{1}<\nu_{2}$ or
$\nu_{2}<\nu_{1}$.

The list of strongly multiplicity free representations consists of (first) fundamental representations
for the series $A_{r}, B_{r}, C_{r}$, the exceptional Lie algebra $G_{2}$ (7-dimensional representation) and
all $A_{1}$ representations.

Fundamental weights of the algebra $A_{r}$ are shorter than its roots, but the projections of $\gf$
weights must be given by linear combinations of the roots of the  special subalgebra $\af$ with integer
coefficients, so the first class of strongly multiplicity free representations does not produce such a
projection.

Nevertheless, the union of diagrams of the two fundamental representations of $A_{2}$ with the root
system of $A_{2}$ produces a weight diagram of $G_{2}$. This case corresponds to splint
$\Delta_{G_{2}}=\varphi_{1}( \Delta_{A_{2}})\cup \varphi_{2}(\Delta_{A_{2}})$ which is connected
with the regular subalgebra $A_{2}\subset G_{2}$ (see \cite{2011arXiv1111.6787L}).

The first fundamental representation of $B_{r}$ immediately gives us the splint $\Delta'=\pi_{B_{r}}\left(
\Delta_{D_{r+1}}\right) = \Delta_{B_{r}}\cup \Delta_{A_{1}+\dots+A_{1}}$ corresponding to the Gelfand-Tsetlin
multiplicity-free branching for the special embedding $\mathfrak{so}(2r+1)\to \mathfrak{so}(2r+2)$.

The length of the first fundamental weight of  $C_{r}$ is less than the length of its short root, so
this case can not correspond to an integer subalgebra.

If there are vectors of a different length in the projection $\Delta'$ on the subalgebra $A_{1}$, such a
projection is not multiplicity free, and $\Delta_{\sfr}$ contains parallel roots. The simplest case
is the special embedding $A_{1}\to A_{2}$ with index 4 where $\Delta_{\sfr}$ is the root system
$BC_{1}$. Such systems do not correspond to semisimple Lie algebras, so we can not speak of a
coincidence of branching coefficients for the reduction $\gf\downarrow \af$ with weight
multiplicities of $\sfr$-representations.

%% The only case of the projection $\Delta'$ to $A_{1}$ with root multiplicities not greater than 2 is
%% special embedding $A_{1}\to A_{2}$ with index 1 given by 
%% 
The weight diagram of the seven-dimensional representation of $G_{2}$ together with the $G_{2}$ root system
form the projection of the root system $B_{3}$ which corresponds to the special embedding 
$G_{2}\to B_{3}$. The branching coefficients in this case coincide with the weight multiplicities of
representations of algebra $\sfr=A_{2}$ with the image of the root system consisting of the short roots of
$G_{2}$.

The complete classification of multiplicity-free irreducible representations was obtained in
\cite{howe1995perspectives} (see also \cite{stembridge2003multiplicity}). It consists of minuscule
and quasi-minuscule representations.

The weight $\mu$ is minuscule if $\left<\mu,\alpha^{\vee}\right>\leq 1$ for all $\alpha\in
\Delta^{+}$ and all the weights of the irreducible representation $L^{\mu}$ lie on the Weyl
group orbit of $\mu$. A weight is quasi-minuscule if $\left<\mu,\alpha^{\vee}\right>\leq 2$ for all
$\alpha\in \Delta^{+}$ and all non-zero weights lie on the single Weyl group orbit.

Minuscule representations are well-studied and very useful in the computations of e.g. tensor products
\cite{stembridge2003multiplicity,stembridge2001computational}. The minuscule representations are
indexed by the weight lattice modulo the root lattice. There is a unique quasi-minuscule
representation that is not minuscule for each simple Lie algebra. The multiplicity of the zero 
weight in quasi-minuscule representation is the number of the short nodes of the Dynkin diagram.

The list of the minuscule representations consists of tensor powers of the vector
representation for the series $A_{r}$; spin representations for the series
$B_{r}$; vector representations for $C_{r}$; vector and two half-spin for $D_{r}$; two
27-dimensional representations for $E_{6}$ and the 56-dimensional representation of $E_{7}$.

Quasi-minuscule representations that are not minuscule are: the adjoint representation of $A_{r}$,
the vector representation for $B_{r}$, the $2r^{2}-r-1$-dimensional representation for $C_{r}$, the adjoint
representations for $D_{r}, E_{6}, E_{7}, E_{8}$, the 26-dimensional representation for $F_{4}$ and
the 7-dimensional of $G_{2}$.

@@

The classification of the multiplicity-free irreducible highest weight modules $L^{\mu}$ in
\cite{howe1995perspectives,stembridge2003multiplicity} consists of the following classes:
\begin{itemize}
\item (1) $\mu$ is minuscule,
\item (2) $\mu$ is quasi-minuscule and Lie algebra has only one short simple root,
\item (3) Lie algebra is $\mathfrak{sp}(6)$ and $\mu=\omega_{1}$, or
\item (4) Lie algebra is $\mathfrak{sl}(n + 1)$ and  $\mu= m\omega_{1}$ or $\mu  = m\omega_{n}$ 
\end{itemize}

We see that the case (1) contains strongly multiplicity free modules of the series $A_{r}$ and the exceptional
Lie algebra $G_{2}$, the class (2) contains strongly multiplicity free modules of $ B_{r}$ and $C_{r}$ ,
strongly multiplicity free modules of $A_{1}$ are included in the class (4).

We do not need to consider the minuscule representations for the simply-laced Lie algebras, since in this
case the length of the weights of the representation $M^{\chi}_{\af}$ is less than the length of the $\af$ roots,
that contradicts to $\af$ being an integer subalgebra. For the series $B_{r}$ we have already
discussed the minuscule spin representation, since it is strongly multiplicity free. The last case in
(1) is the vector representation for $C_{r}$. It is $2r$-dimensional, but the length of its highest weight $\omega_{1}$ is less
than the length of the short root of algebra $C_{r}$. So it can not be a characteristic representation
for the special subalgebra.

%% , so with the dimension of the
%% adjoint representation we get $2r+2r^{2}=2r(r+1)$ which is exactly the dimension of the adjoint
%% representation for the algebra $D_{r+1}$.  
%% But there is no such embedding (?).
%% 
The case (2) consists of the series $B_{r}$ and the exceptional root system $G_{2}$. The embedding $G_{2}\to
B_{3}$ was already discussed. 
 The dimension of the adjoint representation for the series $B_{r}$
is $2r^{2}+r$ and of the quasi-minuscule representation it is $2r+1$, so the dimension of the algebra
$\gf$ should be $(r+1)(2r+1)$ with the $\mathrm{rank}\gf=r+1$. The only solution is $\gf=D_{r+1}$
and we have already seen that this case corresponds to the strongly multiplicity free module and
the Gelfand-Tsetlin basis.

In the case (3) the diagram of the representation contains weights with the length smaller than the
length of the short root, so it cannot be a projection of $\Delta_{\gf}$ to an integer subalgebra.

We need to consider representations of the class (4) only for $m=1,2$ since the root system $\Delta_{\sfr}$
can have roots of at most two different lengths. In the case $m=1$ we get strongly multiplicity free
modules of $A_{r}$ which were discussed above. And for algebra $A_{2}$ and $m=2$ the weight diagram
of the representation $L^{2\omega_{1}}_{A_{2}}$ coincides with the root system of the exceptional Lie algebra
$G_{2}$. This case appears in the special embedding $A_{2}\to B_{3}$. But for $A_{r}, r>2$ there are
no root systems that coincide with the weight diagram of $L^{2\omega_{1}}_{A_{r}}$.

So far we've considered all the cases when $\mathrm{rank}\gf-\mathrm{rank}\af=1$. To complete our
classification of splints we need the classification of all the representations where multiplicities
of all non-zero weights are equal to one. Fortunately, such a classification was obtained in
\cite{plotkin1998visual}. It consists of multiplicity-free representations and adjoint
representations of algebras $A_{r}, B_{r}, C_{r}$, $F_{4}$ and $G_{2}$.

In order to have the embedding $\af\to \gf$ such that the adjoint representation of $\gf$ is
decomposed into two adjoint representations of $\af$ (and possibly several trivial
representations) we need to satisfy the following conditions:
\begin{itemize}
\item Denote by $n_{\gf}$ the total number of roots in $\Delta_{\gf}$ and the projection $\Delta'$.
  Then
  \begin{equation}
    \label{eq:5}
    n_{\gf}=2n_{\af}.
  \end{equation}
\item Denote rank of $\gf$ by shorthand notation $r_{\gf}$. Then the dimension
  \begin{equation}
    \label{eq:6}
    \mathrm{dim}\gf=n_{\gf}+r_{\gf}\geq 2\mathrm{dim}\af=2n_{\af}+2r_{\af}
  \end{equation}

>>>>

Специальные вложения подалгебр детально рассматривались в фундаментальных работах Евгения
Борисовича Дынкина \cite{dynkin1952semisimple,dynkin1952maximal}, где для них использовалось
название ``S-подалгебры'' чтобы отличить от регулярных или ``R-подалгебр'', полученных путем
исключения некоторых корней из корневой системы алгебры. Все регулярные подалгебры нетрудно построить
путем исключения вершин из расширенной диаграммы Дынкина для алгебры, но для специальных подалгебр
такого простого метода нет. Полная классификация для исключительных алгебр Ли была построена недавно
в работе \cite{minchenko2006semisimple}. Алгоритм построения специальных подалгебр реализован в виде
пакета для системы GAP, но построение все равно требует участия человека \cite{de2011constructing}.

Будем предполагать, что алгебра Ли $\gf$ простая. Обозначим подалгебру через $\af$. Соответствующие
подалгебры Картана будем обозначать через $\hfg$ и $\hfa$, и отождествим их с дуальными
пространствами $\hfg^{*}$, $\hfa^{*}$ используя формы Киллинга алгебр  $\gf$ и $\af$.

Для построения вложения  $\af\to\gf$ надо отождествить некоторое представление  $L^{\nu}_{\af}$ с
подпространством в алгебре Ли $\gf$. Затем необходимо проверить, что генераторы  $\af$ могут быть
представлены в виде линейных комбинаций генераторов  $\gf$. После отождествления подалгебры Картана
$\hfa\subset \af$ с дуальным пространством $\hfa^{*}$ при помощи формы Киллинга, корневую систему
$\Delta_{\gf}$ алгебры $\gf$ можно спроектировать на  $\hfa^{*}$ используя выражение
генераторов подалгебры $\hfa$ через генераторы  $\hfg$.

@@

Свойства этой проекции описаны в классических работах  \cite{dynkin1952semisimple,dynkin1952maximal} в
следующих теоремах:

\begin{theorem}\label{dyn0}
  Если представление  $L^{\mu}_{\gf}$ алгебры $\gf$ индуцирует представление $L^{\tilde\mu}_{\af}$
  подалегбры $\af$, то весовая диаграмма $L^{\tilde\mu}_{\af}$ получается из диаграммы
  $L^{\mu}_{\gf}$ ортогональной проекцией $\hfg^{*}$ на $\hfa^{*}$.
  \cite{dynkin1952maximal}. 
\end{theorem}

Применяя эту теорему к присоединенному представлению  $\gf$, видим, что проекция корневой
системы  $\Delta_{\gf}$ на $\hfa^{*}$ оказывается весовой системой некоторого конечномерного но не
обязательно неприводимого представления $\af$. Более того, это приводимое представление должно
содержать присоединенное представление $\af$. Если обозначить проекцию корневой системы через
\begin{equation}
  \label{eq:2}
  \Delta'=\pia\left(\Delta_{\gf}\right),
\end{equation}
то все корни $\af$ будут содержаться в $\Delta'$. То есть система  $\Delta'$ может быть корневой
системой, но она не обязательно является приведенной и может содержать некоторые векторы с
кратностью больше единицы.

Следующие теоремы \cite{dynkin1952semisimple} также описывают свойства  $\Delta'$.

\begin{theorem}\label{dyn1}
  Каждая специальная подалгебра  $\af$ полупростой алгебры Ли  $\gf$ -- целочисленная, то есть
  проекции корней  $\gf$ на $\hfa^{*}$ являются линейными комбинациями простых корней  $\af$ с
  целочисленными коэффициентами \cite{dynkin1952semisimple}
\end{theorem}

\begin{theorem}\label{dyn2}
  Если $\af$  -- полупростая подалгебра полупростой алгебры Ли $\gf$ и генератор $e_{\alpha}$,
  соответствующий корню $\alpha$ алгебры $\af$, представляется в виде линейной комбинации  $e_{\alpha}=\sum_{\beta}
  e_{\beta}$ генераторов $\gf$ соответствующих корням  $\beta$ алгебры $\gf$, тогда все эти корни
  $\beta$ проектируются в  $\alpha$,
  $\pia(\beta)=\alpha$. 
  \cite{dynkin1972semisimple,dynkin1952semisimple}. 
\end{theorem}

Из этих теорем мы видим, что  корневая система $\Delta_{\af}$ вложена метрически (в смысле
определения сплинта) в проекцию  $\Delta'$. Более того, кратности некоторых корней 
$\alpha\in\Delta_{\af}$ в этой проекции  $\Delta'$ больше единицы.

Например, проекция корневой системы $D_{4}$($so(8)$) на корневую систему специальной подалгебры
$A_{2}$($su(3)$) совпадает с корневой системой $G_{2}$, но кратность коротких корней равна $3$.

@@

\section{Сплинты для специальных вложений}
\label{sec:splints-spec-embedd}

@@

Мы хотим перечислить все случаи совпадения коэффициентов ветвления с кратностями весов в
представлении некоторой другой алгебры. Такое совпадение возможно, если проекция $\Delta'$ корневой
системы алгебры $\gf$ допускает сплинт $\Delta'=\varphi_{\af}(\Delta_{\af})\cup
\varphi_{\sfr}(\Delta_{\sfr})$, где  $\varphi_{\af}$ и $\varphi_{\sfr}$ -- вложения соответствующих
корневых систем. Вложение  $\varphi_{\af}$ метрическое и тривиальное. Кроме этого, проекция
сингулярного элемента  $\pia\left(\Psi^{\mu}_{\gf}\right)$ должна раскладываться в линейную
комбинацию сингулярных элементов неприводимых представлений $\sfr$:
\begin{equation}
  \label{eq:4}
  \pia\left(\Psi^{\mu}_{\gf}\right)=\sum_{\nu} \varkappa_{\nu}e^{\nu}\Psi^{\tilde\mu}_{\sfr},
\end{equation}
с целыми коэффициентами  $\varkappa_{\nu}$, где  множество весов  $\nu$, по которому ведется
суммирование, будет определено ниже. Это разложение сингулярного элемента очень похоже на ветвления
орбит группы Вейля, рассмотренные в работах  \cite{larouche2011branching,larouche2009branching}.
Сперва мы перечислим все расщепления проекций корневой системы в объединение простых корневых систем
при специальных вложениях, а затем обсудим разложение сингулярного элемента.

Проекция  $\Delta'$ корневой системы  $\gf$ совпадает с проекцией весовой диаграммы присоединенного
представления алгебры  $\gf$, за исключением нулевого веса. Присоединенное представление  $\gf$
содержит присоединенное представление  $\af$ и раскладывается в прямую сумму
\begin{equation}
  \label{eq:3}
  \mathrm{ad}_{\gf}=\mathrm{ad}_{\af}\oplus M_{\af}^{\chi},
\end{equation}
где  $M^{\chi}_{\af}$ -- не обязательно неприводимое представление $\af$, названное 
``характеристическим'' в работе \cite{dynkin1952semisimple}.

Весовая диаграмма $M^{\chi}_{\af}$ совпадает с корневой системой  $\Delta_{\sfr}$, за исключением
нулевого веса. То есть нам достаточно найти все представления для всех простых алгебр Ли $\af$
такие, что их весовая диаграмма совпадает с некоторой корневой системой после исключения нулевого
веса. Во-первых, заметим, что все веса  $M^{\chi}_{\af}$ должны быть  не меньшей длины чем
кратчайший корень $\af$, так как $\af$ -- целочисленная подалгебра $\gf$ \ref{dyn1}. Во-вторых,
$M^{\chi}_{\af}$ должна содержать веса не более, чем двух разных длин.

Если в корневой системе  $\Delta_{\gf}$ есть корни $\alpha: \alpha\perp \beta,\; \forall \beta\in
\Delta_{\af},$ ортогональные корневой системе подалгебры, то сингулярный элемент не допускает
разложения \ref{eq:4}.  В этом случае элементы формальной алгебры $e^{\nu}$ проекции
$\pia\left(\Psi^{\mu}_{\gf}\right)$ надо снабдить размерностями представлений алгебры $\afb$,
порожденной генераторами, соответствующими ортогональным корням \cite{2010arXiv1007.0318L}. Тогда
правая часть разложения  \ref{eq:4} будет содержать нетривиальные кратности и совпадение
коэффициентов ветвления с кратностями весов в представлениях  $\sfr$ не будет иметь места. В
этом случае можно вывести более сложное соотношение между кратностями весов и коэффициентами
ветвления, но этот вывод лежит за рамками данной работы.

Кратность нулевого веса в присоединенном представлении равна рангу алгебры. Так что при 
$\mathrm{rank}\gf-\mathrm{rank}\af=1$, характеристическое представление  $M^{\chi}_{\af}$ должно
быть свободно от нетривиальных кратностей.  При $\mathrm{rank}\gf-\mathrm{rank}\af>1$ только нулевой
вес может иметь нетривиальную кратность.

Простейшие свободные от кратностей представления называется ``строго свободными от кратностей''
 (``strongly multiplicity free'') \cite{lehrer2006strongly} и представляют собой представления с
 весовыми системами, допускающими строгое упорядочение $\nu_{1}<\nu_{2} \Leftrightarrow
 \nu_{1}=\nu_{2}+n \alpha$, где $n\in
\mathbb{N}, \alpha\in \Delta^{+}_{\af}$ и для всех $\nu_{1},\nu_{2}$ либо $\nu_{1}<\nu_{2}$, либо
$\nu_{2}<\nu_{1}$.

Список строго свободных от кратностей представлений состоит из (первых) фундаментальных
представлений в сериях $A_{r}, B_{r}, C_{r}$, представления исключительной алгебры Ли $G_{2}$
(7-мерного) и всех представлений $A_{1}$.

Фундаментальные веса алгебры $A_{r}$ короче, чем ее корни, но проекции весов $\gf$ должны даваться
линейными комбинациями корней специальной подалгебры  $\af$ с целыми коэффициентами, так что первый
класс строго свободных от кратностей представлений не порождает проекции корневой системы,
допускающей расщепление.

Тем не менее, объединение диаграмм двух фундаментальных представлений  $A_{2}$ с корневой системой
$A_{2}$ образует корневую систему $G_{2}$. Этот случай соответствуют сплинту
$\Delta_{G_{2}}=\varphi_{1}( \Delta_{A_{2}})\cup \varphi_{2}(\Delta_{A_{2}})$, связанному с
регулярной подалгеброй $A_{2}\subset G_{2}$ (подробнее см. \cite{2011arXiv1111.6787L}).

Первое фундаментальное представление  $B_{r}$ немедленно дает сплинт $\Delta'=\pi_{B_{r}}\left(
\Delta_{D_{r+1}}\right) = \Delta_{B_{r}}\cup \Delta_{A_{1}+\dots+A_{1}}$, соответствующий правилу
ветвления Гельфанда-Цетлина с тривиальными коэффициентами для специального вложения
$\mathfrak{so}(2r+1)\to \mathfrak{so}(2r+2)$.

Длина первого фундаментального веса алгебры   $C_{r}$ меньше, чем длина ее короткого корня, так что
этот случай не может порождать сплинт для целочисленной подалгебры.

Далее, если в проекции  $\Delta'$ на подалгебру  $A_{1}$ есть векторы разной длины, то в такой
проекции есть нетривиальные кратности и система $\Delta_{\sfr}$ должна содержать сонаправленные
корни. Простейший пример -- это специальное вложение $A_{1}\to A_{2}$ с индексом 4, где
$\Delta_{\sfr}$ -- это корневая система $BC_{1}$. Такие системы не соответствуют полупростым
алгебрам Ли, так что мы не можем говорить о совпадении коэффициентов ветвления для редукции
$\gf\downarrow \af$ с кратностями весов в представлениях $\sfr$.

%% The only case of the projection $\Delta'$ to $A_{1}$ with root multiplicities not greater than 2 is
%% special embedding $A_{1}\to A_{2}$ with index 1 given by 
%%

Наконец, весовая диаграмма семимерного представления $G_{2}$ вместе с корневой системой $G_{2}$
образует проекцию корневой системы $B_{3}$, возникающую при специальном вложении $G_{2}\to B_{3}$.
Коэффициенты ветвления в этом случае совпадают с кратностями весов в представлениях алгебры $G_{2}$.

Полная классификация представлений, свободных от кратностей, получена в работе
\cite{howe1995perspectives} (см. также \cite{stembridge2003multiplicity}). Она состоит из
микровесовых и квазимикровесовых представлений.

Вес  $\mu$ называется микровесом, если  $\left<\mu,\alpha^{\vee}\right>\leq 1$ для всех $\alpha\in
\Delta^{+}$ и все веса неприводимого представления  $L^{\mu}$ лежат на той же орбите группы Вейля,
что и $\mu$. Вес называется квазимикровесом, если $\left<\mu,\alpha^{\vee}\right>\leq 2$ для всех
$\alpha\in \Delta^{+}$ и все ненулевые веса представления лежат на одной орбите группы Вейля.

Микровесовые представления хорошо изучены и очень полезны при вычислениях, например, тензорных произведений
\cite{stembridge2003multiplicity,stembridge2001computational}. Микровесовые представления нумеруются
элементами весовой решетки, факторизованной по корневой решетке. Для каждой простой алгебры Ли
существует единственное квазимикровесовое представление, не являющееся микровесовым.  Кратность
нулевого веса в квазимикровесовом представлении равна числу коротких вершин в диаграмме Дынкина.

Полный список микровесовых представлений состоит из тензорных степеней векторного представления
алгебр серии $A_{r}$; спинорных представлений алгебр серии $B_{r}$; векторных представлений $C_{r}$;
векторных и полуспиновых представлений $D_{r}$; двух 
27-мерных представлений алгебры $E_{6}$ и  56-мерного представления $E_{7}$.

Квазимикровесовые представления, не являющиеся микровесовыми -- это присоединенное представление
$A_{r}$, векторное представление $B_{r}$, представление размерности $2r^{2}-r-1$ алгебры $C_{r}$,
присоединенные представления алгебр $D_{r}, E_{6}, E_{7}, E_{8}$, 26-мерное представление $F_{4}$ и
7-мерное представление $G_{2}$.

@@

Классификация неприводимых представлений  $L^{\mu}$, свободных от кратностей, в работе
\cite{howe1995perspectives,stembridge2003multiplicity} состоит из следующих классов:
\begin{itemize}
\item (1) $\mu$ -- микровес
\item (2) $\mu$ -- квазимикровес и в алгебре Ли только один короткий простой корень,
\item (3) алгебра Ли $\mathfrak{sp}(6)$ и $\mu=\omega_{1}$, и
\item (4) алгебра Ли $\mathfrak{sl}(n + 1)$ и  $\mu= m\omega_{1}$ или $\mu  = m\omega_{n}$ 
\end{itemize}

Как видно, класс (1) содержит строго свободные от кратностей представления серии  $A_{r}$ и
исключительной алгебры Ли $G_{2}$, класс  (2) содержит строго свободные от кратностей представления
$ B_{r}$ и $C_{r}$ , а строго свободные от кратностей представления $A_{1}$ входят в класс (4).

Нам не надо рассматривать микровесовые представления алгебр $A$, $D$, $E$, так как в этом случае
длина весов представления  $M^{\chi}_{\af}$ меньше длины корней подалгебры  $\af$, что противоречит
целочисленности подалгебры  $\af$. Микровесовые спиновые представления алгебры серии $B_{r}$ уже
обсуждались, так как они являются строго свободными от кратностей. Последний случай в классе (1) --
это векторное представление  $C_{r}$. Оно  $2r$-мерно, но длина старшего веса $\omega_{1}$ меньше
длины короткого корня алгебры $C_{r}$, так что это представление не может быть характеристическим
для специальной подалгебры.

%% , so with the dimension of the
%% adjoint representation we get $2r+2r^{2}=2r(r+1)$ which is exactly the dimension of the adjoint
%% representation for the algebra $D_{r+1}$.  
%% But there is no such embedding (?).
%%

Класс (2) состоит из серии  $B_{r}$ и исключительной корневой системы $G_{2}$. Вложение  $G_{2}\to
B_{3}$ уже описано выше. Размерность присоединенного представления для серии $B_{r}$ равна
$2r^{2}+r$, а размерность квазимикровесового представления равна $2r+1$, таким образом размерность
алгебры 
$\gf$ должна быть $(r+1)(2r+1)$ при ранге $\mathrm{rank}\gf=r+1$. Единственное решение,
удовлетворяющее этим условиям -- это  $\gf=D_{r+1}$ и мы уже видели, что этот случай соответствует
строго свободным от кратностей представлениям и базису Гельфанда-Цетлина.

В классе  (3) диаграмма представления содержит веса с длиной меньше длины короткого корня, так что
она не может быть проекцией  $\Delta_{\gf}$ на целочисленную подалгебру.

Нам следует рассмотреть представления класса  (4) только при $m=1,2$, так как корневая система
$\Delta_{\sfr}$ может содержать корни не более чем двух разных длин. В случае case $m=1$ мы снова
получаем строго свободные от кратностей модули $A_{r}$, которые уже обсуждались выше. А для алгебры
$A_{2}$ и $m=2$ весовая диаграмма представления $L^{2\omega_{1}}_{A_{2}}$ совпадает с корневой
системой исключительной алгебры Ли
$G_{2}$. Этот случай возникает в специальном вложении  $A_{2}\to B_{3}$. Для $A_{r}, r>2$ не
существует корневых систем, которые совпадают с весовой диаграммой $L^{2\omega_{1}}_{A_{r}}$.

Мы рассмотрели все случаи, когда $\mathrm{rank}\gf-\mathrm{rank}\af=1$. Чтобы завершить нашу
классификацию сплинтов, нам нужна классификация всех представлений, в которых кратности ненулевых
весов равны единице. Такая классификация была представлена в работе
\cite{plotkin1998visual}. Она состоит из свободных от кратностей представлений и присоединенных
представлений алгебр  $A_{r}, B_{r}, C_{r}$, $F_{4}$ и $G_{2}$.

Чтобы существовало вложение $\af\to \gf$, такое что присоединенное представление  $\gf$
раскладывается в два присоединенных представления $\af$ (и, возможно, несколько тривиальных
представлений), должны быть выполнены следующие условия:
\begin{itemize}
\item Обозначим через $n_{\gf}$ общее число корней в  $\Delta_{\gf}$ и проекции $\Delta'$.
  Тогда
  \begin{equation}
    \label{eq:5}
    n_{\gf}=2n_{\af}.
  \end{equation}
\item Обозначим ранг $\gf$ для краткости через $r_{\gf}$. Тогда размерность
  \begin{equation}
    \label{eq:6}
    \mathrm{dim}\gf=n_{\gf}+r_{\gf}\geq 2\mathrm{dim}\af=2n_{\af}+2r_{\af}
  \end{equation}
